Collision rates of planetesimals near mean-motion resonances

### 2021-02-23 refereeResponse2.txt

1) In page 7 is said — "We assume that the eccentricity and inclination dispersions are coupled (Ida et al. 1993) and use the eccentricity dispersion as an independent variable.” — It is strange to think on the eccentricity as an independent parameter once it is coupled with the inclination. Maybe change it to “free parameter”?

    "independent variable" has been changed to "free parameter" here.

2) The explanation for the changes in <\sigma v> in the caption of Figure 5 seems incorrect. Actually, their are correct, but what the authors call lef-hand points should be right-hand points and vice-versa. The correct explanation (relation left- right-points) is given in the main text (page 7 - lines 20—25; “… Near the 3:1… drive it higher). Please, review it.

    Fixed

3) Page 7 — “However, one can safely ignore this process if the peak radial drift timescale is longer than the timescale for collisions between these bodies.” — I think safely is too strong! Please, consider changing to “can safely” or “may”.

    "can safely" has been changed to "may be able to".

4) Wallace et al. assume values for \Sigma_0 and \rho_0 in the minimum solar nebular that are not conventional. I understand they are comparing their model with Richardson et al. (2000), and thus, they are using the same values from that work. But, please, consider given the reader a more physical explanation. For example:

4.1) In section 2.4 the authors use 4e-10 g/cm^3 for minimum solar nebula density at 3 au. It is rather confusing why this value was chosen. The density of the nebula in Hayashi (1981) is given by \rho_r = \rho_0*[r^(-2.75)] — eq 2.7 assuming z=0 (mid-plane) and \rho_0 = 1.4e-9 g/cm^3. At 3 au this will result \rho(r=3au) ~ 2.86e-11 g/cm^3 and not 4e-10 g/cm^3. Are the authors considering a different gas temperature profile (what is the flaring relation — did I miss that in the text)? Please, consider reviewing this value or explaining why not using the value from quoted reference. The same value is used in section 5 to explain the fate of the dust.

   The referee is correct that we should have used \rho(r=3au) ~ 2.86e-11 g/cm^3. The value in the paper draft comes from
   a mistake in our calculations, where we accidentally scaled \rho_0 as a surface density, rather than as a volume density.
   It appears that this mistake was only made when calculating a value for \rho(r=3au) to display in the paper and does not appear in any of our analysis code. Therefore, the gas drag timescale in section 2.4, the mean free path in section 5, and the radial drift timescales in section 5 remain the same.

4.2) In section 3.2 it is once again not clear how Wallace et al have chosen the minimum nebula value for \Simga_0. In Hayashi (1981) the surface density is defined as \Sigma_0 = 7.1 g/cm^2 for rock bodies only, 30 g/cm^2 for rock+ice bodies, and 1700 g/cm^2 for gas. Where does \Sigma_0 = 10  g/cm^2 comes from? Please, consider explaining this.

  Following other papers (which cite Hayashi 1981 and then use \Sigma_0 = 10 g/cm^2), we have now changed the sentence before equation 9 to read "The planetesimal disk loosely follows a minimum-mass solar nebula profile"

5) I would once again call the attention for quoting figures from other works. Instead of “(see figure 3c and 3f)” use (see figure 3c and 3f from their work), or (see their figures 3c and 3f). It sounds like a useless change, but I caught myself once again like, wait, what, are there panels 3(a, b, c, d, e, and f) in this work? Same for (figure 3 a). All this is in section 4.1.

  "see figure" has been changed to "see their figure(s)"

### 2021-02-23 main.tex
Final edits before acceptance to journal

\noindent where $m$ and $s$ are the mass and radius of the planetesimal. $C_{D}$ is a drag coefficient which is of order unity, $
\rho_{g}$ is the local density of the gas and $v_{g}$ is the headwind velocity of the gas experienced by the planetesimal. At 3 au, the gaseous component of the solar nebula has a
density of $4 \times 10^{-10}$ g cm$^{-3}$ and the typical headwind experienced by a body on a Keplerian orbit is $\sim 5,000$ cm s$^{-1}$ \citep{1981PThPS..70...35H}. For a 100 km 
body with a density of 2 g cm$^{-3}$ (assuming a mixture of ice and rock) the stopping timescale is on the order of 1 Myr. This is much longer than the timescales associated with 
secular forcing and libration due to mean-motion resonances, as discussed above, but shorter than the typical protoplanetary disk lifetime and potentially shorter the planetesimal 
formation timescale. For this reason, we do not model the effects of gas drag in the simulations, although we consider its effects when constructing initial conditions, which is discussed in 
the next section.

@@

In total, five simulations are run, which are listed in table \ref{tab:sims}. The second is a ``nominal'' case, in which the perturbing planet's mass and eccentricity are set to that of Jupiter's. 
In the other four cases, the mass or eccentricity is altered by a factor of two from the nominal value. In all cases, the perturbing giant is placed on a 5.2 au orbit around a 1 
$M_{\odot}$ star. The planetesimal disc extends from 2.2 to 3.8 au, which covers the two most prominent mean-motion resonances with the giant planet, the 2:1 at 3.27 au and the 3:1 at 
2.5 au. The planetesimal disc follows a minimum-mass solar nebula surface density profile \citep{1981PThPS..70...35H}

@@

We begin by examining simulations e1m2, e2m2 and e3m2. The positions of the planetesimals in the x-y plane after 5,000 years of integration are 
shown in figure \ref{fig:xy}. In all cases, the coordinate system is rotated so that the giant planet lies at $\theta = 0$. The longitude of perihelion of 
the planet is shown by the dashed line. Resonances with the perturbing planet are visible as nonaxisymmetric gaps in this figure. Upon close inspection of a 
series of simulation snapshots, these gaps appear to follow the planet in its orbit, rather than aligning themselves with the longitude of perihelion. A 
similar substructure, which is most noticeable near the 2:1 MMR, reveals itself in \citet{2000Icar..143...45R} (see figure 3c and 3f) and 
\citet{2016ApJ...818..159T} (figure 3a). This structure is also present in the \citet{2017ApJ...850..103B} simulations, although it was not reported at the 
time (see Appendix \ref{sec:boley_plot}). It is worth noting that both \citet{2000Icar..143...45R} and \citet{2017ApJ...850..103B} started with a completely cold planetesimal disc. 
Thus, the presence of these features seems robust to the choice of initial conditions.

@@

In figure \ref{fig:gf}, we show the effect that the 2:1 MMR and 3:1 MMR have on the local average collision rate. Each pair of points connected by a line represents the average 
collision rate before and after the effects of the mean-motion resonances develop. Here, the relative velocity between bodies is calculated by measuring the eccentricity and inclination 
dispersion near each resonance and using equation \ref{eq:eccincvel}. We assume that the eccentricity and inclination dispersions are coupled \citep{1993MNRAS.263..875I} and use the 
eccentricity dispersion as an independent variable. Because both the strength of secular forcing and the effects of the resonant perturbations are different near the 3:1 and 2:1 MMRs, the 
planetesimal population at each of these locations in the disk experiences a qualitatively different change in the collision rate. Near the 3:1 MMR, the unperturbed collision rate is well 
above the minimum value $\left< \sigma v \right>_{0}$ and the dynamical heating introduced by the resonance acts to suppress it further. Near the 2:1 MMR, the collision rate starts out 
much closer to the minimum value and the additional perturbations instead act to drive it higher.

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{figures/gf_plot.png}
    \caption{The collision rate (given by equation \ref{eq:saf}) of bodies in the vicinity of the 3:1 (left) and 2:1 (right) MMRs, relative to the minimum value, as a function of the local 
    eccentricity dispersion. The pairs of points connected by lines show the values of the perturbed (left-hand points in each subplot) and unperturbed (right-hand points) collision rates in 
    the e1m2, e2m2 and e3m2 simulations. The unperturbed collision rate, along with the amount of dynamical heating that the planetesimal population experiences determines whether 
    the MMR acts to suppress or enhance the collision rate.\label{fig:gf}}
\end{center}
\end{figure*}

@@

We first provide a justification for assumption (1), which is as follows: At 3 au in the protosolar nebula, the mean free path of a gas particle is $\sim$ 50 cm (assuming a composition 
of pure hydrogen with a local volume density of $\rho_{g} = 4 \times 10^{-10}$ g cm$^{-3}$). This places any dust grains in the Epstein drag regime, with a stopping timescale given by

@@

Assumption (2) is less straightforward to justify and may not be reasonable in all cases. Although the spatial redistribution of the full-sized planetesimals and mm-sized dust grains 
is insignificant, the radial drift forces experienced by intermediate-sized bodies is much greater. Collisions between planetesimals do not immediately convert the mass to dust grains. 
Rather, the debris will undergo a collisional cascade process, which
gradually grinds the resulting bodies down to smaller sizes. A proper
treatment of the collisional cascade process is 
quite complicated and is well beyond the scope of this work. However, one can safely ignore this process if the peak radial drift timescale is longer than the timescale for collisions 
between these bodies.

>>>>

\noindent where $m$ and $s$ are the mass and radius of the planetesimal. $C_{D}$ is a drag coefficient which is of order unity, $
\rho_{g}$ is the local density of the gas and $v_{g}$ is the headwind velocity of the gas experienced by the planetesimal. At 3 au, the gaseous component of the solar nebula has a
density of $3 \times 10^{-11}$ g cm$^{-3}$ and the typical headwind experienced by a body on a Keplerian orbit is $\sim 5,000$ cm s$^{-1}$ \citep{1981PThPS..70...35H}. For a 100 km 
body with a density of 2 g cm$^{-3}$ (assuming a mixture of ice and rock) the stopping timescale is about 1 Myr. This is much longer than the timescales associated with 
secular forcing and libration due to mean-motion resonances, as discussed above, but shorter than the typical protoplanetary disk lifetime and potentially shorter the planetesimal 
formation timescale. For this reason, we do not model the effects of gas drag in the simulations, although we consider its effects when constructing initial conditions, which is discussed in 
the next section.

@@

In total, five simulations are run, which are listed in table \ref{tab:sims}. The second is a ``nominal'' case, in which the perturbing planet's mass and eccentricity are set to that of Jupiter's. 
In the other four cases, the mass or eccentricity is altered by a factor of two from the nominal value. In all cases, the perturbing giant is placed on a 5.2 au orbit around a 1 
$M_{\odot}$ star. The planetesimal disc extends from 2.2 to 3.8 au, which covers the two most prominent mean-motion resonances with the giant planet, the 2:1 at 3.27 au and the 3:1 at 
2.5 au. The planetesimal disc loosely follows a minimum-mass solar nebula surface density profile \citep{1981PThPS..70...35H}

@@

We begin by examining simulations e1m2, e2m2 and e3m2. The positions of the planetesimals in the x-y plane after 5,000 years of integration are 
shown in figure \ref{fig:xy}. In all cases, the coordinate system is rotated so that the giant planet lies at $\theta = 0$. The longitude of perihelion of 
the planet is shown by the dashed line. Resonances with the perturbing planet are visible as nonaxisymmetric gaps in this figure. Upon close inspection of a 
series of simulation snapshots, these gaps appear to follow the planet in its orbit, rather than aligning themselves with the longitude of perihelion. A 
similar substructure, which is most noticeable near the 2:1 MMR, reveals itself in \citet{2000Icar..143...45R} (see their figures 3c and 3f) and 
\citet{2016ApJ...818..159T} (see their figure 3a). This structure is also present in the \citet{2017ApJ...850..103B} simulations, although it was not reported at the 
time (see Appendix \ref{sec:boley_plot}). It is worth noting that both \citet{2000Icar..143...45R} and \citet{2017ApJ...850..103B} started with a completely cold planetesimal disc. 
Thus, the presence of these features seems robust to the choice of initial conditions.

@@

In figure \ref{fig:gf}, we show the effect that the 2:1 MMR and 3:1 MMR have on the local average collision rate. Each pair of points connected by a line represents the average 
collision rate before and after the effects of the mean-motion resonances develop. Here, the relative velocity between bodies is calculated by measuring the eccentricity and inclination 
dispersion near each resonance and using equation \ref{eq:eccincvel}. We assume that the eccentricity and inclination dispersions are coupled \citep{1993MNRAS.263..875I} and use the 
eccentricity dispersion as a free parameter. Because both the strength of secular forcing and the effects of the resonant perturbations are different near the 3:1 and 2:1 MMRs, the 
planetesimal population at each of these locations in the disk experiences a qualitatively different change in the collision rate. Near the 3:1 MMR, the unperturbed collision rate is well 
above the minimum value $\left< \sigma v \right>_{0}$ and the dynamical heating introduced by the resonance acts to suppress it further. Near the 2:1 MMR, the collision rate starts out 
much closer to the minimum value and the additional perturbations instead act to drive it higher.

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{figures/gf_plot.png}
    \caption{The collision rate (given by equation \ref{eq:saf}) of bodies in the vicinity of the 3:1 (left) and 2:1 (right) MMRs, relative to the minimum value, as a function of the local 
    eccentricity dispersion. The pairs of points connected by lines show the values of the unperturbed (left-hand points in each subplot) and perturbed (right-hand points in each subplot) collision rates in 
    the e1m2, e2m2 and e3m2 simulations. The unperturbed collision rate, along with the amount of dynamical heating that the planetesimal population experiences determines whether 
    the MMR acts to suppress or enhance the collision rate.\label{fig:gf}}
\end{center}
\end{figure*}

@@

We first provide a justification for assumption (1), which is as follows: At 3 au in the protosolar nebula, the mean free path of a gas particle is $\sim$ 50 cm (assuming a composition 
of pure hydrogen with a local volume density of $\rho_{g} = 3 \times 10^{-11}$ g cm$^{-3}$). This places any dust grains in the Epstein drag regime, with a stopping timescale given by

@@

Assumption (2) is less straightforward to justify and may not be reasonable in all cases. Although the spatial redistribution of the full-sized planetesimals and mm-sized dust grains 
is insignificant, the radial drift forces experienced by intermediate-sized bodies is much greater. Collisions between planetesimals do not immediately convert the mass to dust grains. 
Rather, the debris will undergo a collisional cascade process, which
gradually grinds the resulting bodies down to smaller sizes. A proper
treatment of the collisional cascade process is 
quite complicated and is well beyond the scope of this work. However, one may be able to ignore this process if the peak radial drift timescale is longer than the timescale for collisions 
between these bodies.
